\LARGE \bf Face Recognition: Too Bias, or Not Too Bias?

### 2020-02-15 sections/section3.tex
updated supplemental citations
cropped margins from PDFs
Updates from Overleaf

\subsection{The data}
Problems of bias in \gls{fr} motivated the design of \gls{bfw}. The data evenly represent various subgroups partitioned by demographics. Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit. \gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increased in the number subjects per subgroup, with many more pairs (Table~\ref{tab:ethnic-splits}).

@@

\vspace{1mm}
\noindent\textbf{Detecting faces.} Faces were detected using \gls{mtcnn}~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets. Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set. The second set were faces aligned and cropped for Arcface~\cite{deng2019arcface} (see the next step). Also, coordinates of the BB and the five landmarks per MTCNN were stored as part of the static, raw data. Samples with multiple face detections had the BB area times the confidence score of the MTCNN to determine the instance most likely to be the face, with others set aside and labeled \textit{miss-detection}.

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[trim=7mm 0.1in 5mm 0in,clip,width=\linewidth]{images/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed lined shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the number of match errors, so closer a curve is to the bottom better.}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\vspace{-5mm}
\end{figure}

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Arcface~\cite{deng2019arcface}. A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$. Mathematically, the $j^{th}$ subject with $N_j$ faces is removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, $\theta=0.2$ set manually, and an ordered list of scores. This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. The faces were first ordered from most-to-least confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}. 
% Each subject underwent two passes by different annotators.

%  by flagging \emph{imposters}

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\centering
	\centering
	\includegraphics[trim=0.3in 0.1in 0.2in 0.2in,clip,angle=-0,origin=c,width=1\linewidth]{images/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} The scores for \emph{imposters} have medians about 0.3 but with variation in upper percentiles, while \emph{genuine} pairs vary in both (\eg \gls{af} has more area of overlap). A varying threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model} 
\end{figure}

\vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectfully. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subjects is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

@@

\subsection{Human Assessment}\label{subsec:human-assessment}
We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians. To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C). The purpose was to the minimize variability by only analyzing the subsets of the broader groups of whites and Asians.

>>>>

\subsection{The data}
Problems of bias in \gls{fr} motivated the design of \gls{bfw}. The data evenly represent various subgroups partitioned by demographics. Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit. \gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number subjects per subgroup, and with many more pairs (Table~\ref{tab:ethnic-splits}).

@@

\vspace{1mm}
\noindent\textbf{Detecting faces.} Faces were detected using \gls{mtcnn}~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets. Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set. The second set were faces aligned and cropped for Sphereface~\cite{liu2017sphereface} (see the next step). Also, coordinates of the BB and the five landmarks from \gls{mtcnn} were stored as part of the static, raw data. For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed lined shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the number of match errors, so closer a curve is to the bottom better.}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\end{figure}

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}. A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$. Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, fell below a threshold of $\theta=0.2$ (set manually). This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\centering
	\centering
	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} The scores for \emph{imposters} have medians about 0.3 but with variation in upper percentiles, while \emph{genuine} pairs vary in both (\eg \gls{af} has more area of overlap). A varying threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model}

\end{figure}

% \vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectfully. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subjects is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

@@

\subsection{Human assessment}\label{subsec:human-assessment}
We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians. To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C). The purpose was to the minimize variability by only analyzing the subsets of the broader groups of whites and Asians.

### 2020-02-15 supplemental.tex
Updates from Overleaf

%
% \newpage
\onecolumn
\renewcommand{\thesection}{\alph{section}}
\glsresetall
\setcounter{section}{0}

\section{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exists across different models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
\begin{figure}[h!]
\vspace{-2mm}
    \centering
    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{figures/SDM.pdf}\\
    \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
    \vspace{-3mm}
\end{figure}

Sample faces per subgroup of the \gls{bfw} dataset are shown in Fig.~\ref{fig:montage:app}.
\begin{figure}[h!]
\vspace{-2mm}
    \centering
    \includegraphics[width=.8\linewidth]{figures/facemontage.pdf}
    \caption{\textbf{Sample of \gls{bfw}}. Each row depicts a different gender, \gls{f} (top) and \gls{m} (bottom). Columns are grouped by ethnicity (\ie \gls{a}, \gls{b}, \gls{i}, and \gls{w}, respectfully).}
    \label{fig:montage:app}
    \vspace{-3mm}
\end{figure}

\begin{thebibliography}{widest entry}
 \bibitem[1]{Cao18} Cao, Qiong, Li Shen, Weidi Xie, Omkar M. Parkhi, and Andrew Zisserman. "Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face \& Gesture Recognition.} 2018.
  \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun. "Deep residual learning for image recognition." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
 \bibitem[3]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. "Squeeze-and-excitation networks." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
\end{thebibliography}

>>>>

%
\newpage
\onecolumn
% \begin{minipage*}[\linewidth]{width of the minipage}
\renewcommand{\thesection}{\alph{section}}
\glsresetall
\setcounter{section}{0}

\section{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst). 
% ArcFace (not shown) shown similar results. We would have to cite + will incorporate later .... a blank statement and sphereface is comparable ... up to you i do not care
% \begin{figure}[h!]
% % \vspace{-2mm}
%     \centering
%     \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{figures/SDM.pdf}\\
%     \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
%     \vspace{-3mm}
% \end{figure}

\begin{figure}[h!]
\vspace{-1mm}
    \centering
    \begin{subfig}
    \includegraphics[width=.3\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
    \end{subfig}
    \begin{subfig}
    \includegraphics[width=.28\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
    \end{subfig}
        \begin{subfig}
    \includegraphics[width=.28\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
    \end{subfig}
    \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
    \vspace{-1mm}
\end{figure}

Sample faces per subgroup of the \gls{bfw} dataset are shown in Fig.~\ref{fig:montage:app}.
\begin{figure}[h!]
\vspace{-2mm}
    \centering
    \includegraphics[width=.75\linewidth]{figures/facemontage.pdf}
    \caption{\textbf{Sample of \gls{bfw}}. Each row depicts a different gender, \gls{f} (top) and \gls{m} (bottom). Columns are grouped by ethnicity (\ie \gls{a}, \gls{b}, \gls{i}, and \gls{w}, respectfully).}
    \label{fig:montage:app}
    \vspace{-1mm}
\end{figure}
{
\scriptsize
\begin{thebibliography}{widest entry}
 \bibitem[1]{Cao18} Cao, Qiong, Li Shen, Weidi Xie, Omkar M. Parkhi, and Andrew Zisserman. "Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face \& Gesture Recognition.} 2018.
  \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun. "Deep residual learning for image recognition." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
 \bibitem[3]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. "Squeeze-and-excitation networks." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
\end{thebibliography}
}

% \end{minipage*}
